Approximation as Differentiation notes

### 2025-03-08 notes/sec/grothendieck.tex
Ok can make it work if I can precompose family with (pairing with x).
Type index on eval is a bit painful, but keep it for consistency for now..
Closer.
Clarify f as morphism from constant family.
'precomposition with reindexing function' operation on families of morphism is better understood as the action of the functor induced by the reindexing function.
Slightly more forgiving indexing on eval.
Fix index.
Just enough notation to state proposition that copairing distributes over parallel composition.
Fixing up right-hand commutative diagram.
Corrections.
Correction. Finally getting there.
Ba

\begin{enumerate}
\item $[X, Y]$ to be the object in $\Fam(C)$ with $\idx{[X,Y]} = \Fam(C)(X,Y)$ and $\faml{[X, Y]}(f) =
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
\begin{itemize}
\item $\idx{\lambda}_{X,Y,Z}(f)(x) = g$ where $\idx{g} = \idx{f} \comp (x, -): \idx{Y} \to \idx{Z}$ and
$\faml{g} = \faml{f} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
\faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\faml{f} \comp \inj_{\faml{X}_x}:
\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f}$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$)
\end{itemize}
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
\begin{itemize}
\item $\idx{\eval}_{X,Y}(f, x) = \idx{f}(x)$
\item $\faml{\eval}_{X,Y}(f, x) = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
\end{itemize}
using $\coprodM{-}{-}$ to denote coproduct of morphisms.
\end{enumerate}
Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have the following.
\begin{enumerate}
\item The diagram on the left in $\Fam(C)$ commutes; it suffices to show that the diagram on the right in $C$
commutes for every $(x, y) \in \idx{X} \times \idx{Y}$:

\begin{center}
\begin{tikzcd}
   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
\end{tikzcd}
\hspace{1cm}
\begin{tikzcd}
   \prod_{\idx{Y}}\reindex{\faml{Z}}{y' \mapsto \idx{f}(x,y')} \biprod \faml{Y}_y \arrow[dr,
   "\coprodM{\eval_{\prod_{\idx{Y}}}}{\faml{f}_x}"] \\
   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "\lambda_{\prod_{\idx{Y}}}(y'\;\mapsto\;\faml{f}_{x, y'}\;\comp\;
\inj_{\faml{X}_x}) \biprod \id_{\faml{Y}_y}"] \arrow[r,
   "\faml{f}_{x,y}"'] & \faml{Z}_{\idx{f}(x,y)}
\end{tikzcd}
\end{center}

>>>>

\begin{enumerate}
\item $[X, Y]$ to be the object in $\Fam(C)$ with $\idx{[X,Y]} = \Fam(C)(X,Y)$ and $\faml{[X, Y]}(f) =
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
\begin{itemize}
\item $\idx{\lambda}_{X,Y,Z}(f)(x) = g$ where $\idx{g} = \idx{f} \comp (x, -): \idx{Y} \to \idx{Z}$ and
$\faml{g} = \reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
\faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{X}_x}:
\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{\idx{f}})$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\end{itemize}
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
\begin{itemize}
\item $\idx{\eval}_{X,Y}(f, x) = \idx{f}(x)$
\item $\faml{\eval}_{X,Y}(f, x) = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
\end{itemize}
using $\coprodM{-}{-}$ to denote coproduct of morphisms.
\end{enumerate}
Then for any $f: X \times Y \to Z$ in $\Fam(C)$:
\begin{enumerate}
\item Suppose $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the left below commutes by the universal property
of the set-indexed product; the commutativity of the diagram on the right is trivial.

\begin{center}
\begin{tikzcd}
   \prod_{\idx{Y}}\reindex{\faml{Z}}{\idx{f} \comp (x,-)} \arrow[dr, "\eval_{\prod_{y}}"] \\
   \faml{X}_x \arrow[u, "\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
\inj_{\faml{X}_x})"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}"'] & \faml{Z}_{\idx{f}(x,y)}
\end{tikzcd}
\hspace{1cm}
\begin{tikzcd}[column sep=5em]
   \faml{Y}_y \arrow[dr,
   "\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}"] \\
   \faml{Y}_y \arrow[u, -, double, "\id_{\faml{Y}_y}"] \arrow[r, "\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}"'] & \faml{Z}_{\idx{f}(x,y)}
\end{tikzcd}
\end{center}

The diagram on the left below then commutes by Proposition~\ref{prop:biproduct:prod-coprod}, with
$\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}} =
\faml{f}_{(x,y)}$ by the uniqueness of the coproduct of morphisms. Then the diagram on the right in $\Fam(C)$
commutes by the definitions of $\lambda_{X,Y,Z}, \eval_{Y,Z}$ and $[Y,Z]$.

\begin{center}
\begin{tikzcd}[column sep=5.5em]
   \prod_{\idx{Y}}\reindex{\faml{Z}}{\idx{f} \comp (x,-)} \biprod \faml{Y}_y \arrow[dr,
   "\coprodM{\eval_{\prod_{y}}}{\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}}"] \\
   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "(\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
\inj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r, "\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}}"'] & \faml{Z}_{\idx{f}(x,y)}
\end{tikzcd}
\hspace{1cm}
\begin{tikzcd}
   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
\end{tikzcd}
\end{center}
